---
title       : Canada's Gross Domestic Product (GDP) estimator
subtitle    : An homework assignment for the Developing Data Products course
author      : Etienne Laurin
job         : Student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}

--- .class #id 

## Overview

This is an overview of the application "[Canada's Gross Domestic Product (GDP)](https://etiennelaurin.shinyapps.io/CanadaGDP/)".

To run it, follow the link above and select a year and a territory.

In the right panel, you will see the estimate in million chained (2007) dollars.

--- .class #id 

## Data and Methodology

Origin of the data:
* Statistics Canada, Canada's official federal statistics agency
* Based on the GDP data from 1981 to 2012
* Public data
* CANSIM table 384-0038

Predictors use:
* Province or territory
* Year
 
Method:
* Generalized linear model (GLM)

--- .class #id 

## Purpose

* GDP growth is an indicator of economic prosperity

```{r, echo=FALSE, cache=TRUE, results='hide'}
library(caret)

# Download the file
tf = tempfile()
download.file("http://www20.statcan.gc.ca/tables-tableaux/cansim/csv/03840038-eng.zip", tf)
fname = unzip(tf, list=TRUE)$Name[1]

# Read the data and subset only the data needed
gdp <- read.table(unzip(tf, files=fname), header=TRUE, sep=",", dec = ".", na.strings ="..")
gdp <- subset(gdp[c("Ref_Date","GEO","Value")], gdp$PRI == "Chained (2007) dollars" & gdp$EST == "Final consumption expenditure (x 1,000,000)" & !(is.na(gdp$Value)))

# Split in a training and testing set
set.seed(50)
inTrain = createDataPartition(gdp$Value, p=0.8, list=TRUE)[[1]]
training = gdp[ inTrain,]
testing = gdp[-inTrain,]

# Create a model
model <- train(Value ~ ., data=gdp, method="glm")
predictions <- predict(model, newdata=testing)

# Prediction function
nd <- data.frame(2016, "Quebec")
names(nd) <- c("Ref_Date","GEO")
result <- predict(model, nd)
```

For example, in the year 2016, ceteris paribus, Quebec's GDP should be approximately `r paste(round(result))` million chained (2007) dollars.

--- .class #id 

## Accuracy

```{r, echo=FALSE, cache=TRUE, results='hide'}
pr <- postResample(predictions, testing$Value)
```

Root mean squared error:
`r pr[[1]]`

R squared:
`r pr[[2]]`
